## Approaches to Classification

### Study Goals

*Theoretical (T)*

- Learn how generative model works
- Get familiar with LDA, QDA and naive Bayes classifier
- Generative vs. discriminative approach

*Practical (P)*

- Get to know how to fit a generative model in `mlr`

### Preparation

1.  Watch the following video  (sorry, rather low volume...):
    <center>
    ![](https://www.youtube.com/watch?v=hXQq-lryqtE&list=PLMyWaJl2LoXyhFvMjtbBGs0Pi8khHbKm3&index=4&t=0s){width="75%"}
    </center>

### Exercises
  
#### *(T)* Quiz
  
```{r knn-quiz, echo=FALSE}
question("Which statements are false?",
  answer("In the LDA, each class density is modeled as a multivariate Gaussian with unequal covariance.", correct = TRUE),
  answer("In the QDA, each class density is modeled as a multivariate Gaussian with equal covariance", correct = TRUE),
  answer("In the Naive Bayes classifier, covariance matrices can differ over both classes but have to be diagonal.")
)
```

#### *(P)* Qualitative comparison of LDA, QDA, and naive Bayes

In this demo we want to compare LDA, QDA, and naive Bayes by looking at the decision boundaries. Use `plotLearnerPrediction()` to visualize the decision boundaries for the features `Sepal.Length` and `Sepal.Width`. You can use the build in iris task `iris.task` of `mlr`. Store the plot for each model and plot them among each other with `gird.arrange()` from the `gridExtra` package:

```{r generative-approaches, exercise=TRUE, exercise.lines=8, fig.height=8}
set.seed(123)
gg1 = plotLearnerPrediction(..., task = iris.task, features = ...)
gg2 = plotLearnerPrediction(..., task = iris.task, features = ...)
gg3 = plotLearnerPrediction(..., task = iris.task, features = ...)

grid.arrange(..., ncol = ...)
```

```{r generative-approaches-hint-1}
# The mlr learner are:
"classif.lda"
"classif.qda"
"classif.naiveBayes"
# You can directly use the string instead of a learner.
```

```{r generative-approaches-hint-2}
# You can specify the used features in 'plotLearnerPrediction()' by setting the features argument: 
plotLearnerPrediction("classif.lda", task = iris.task, features = c("Sepal.Length", "Sepal.Width"))
```

```{r generative-approaches-hint-3}
# The plots can be plotted side by side by passing the three ggplots as comma separated arguments: 
grid.arrange(gg1, gg2, gg3, ncol = 1)
```

```{r generative-approaches-solution}
gg1 = plotLearnerPrediction("classif.lda", task = iris.task, features = c("Sepal.Length", "Sepal.Width"))
gg2 = plotLearnerPrediction("classif.qda", task = iris.task, features = c("Sepal.Length", "Sepal.Width"))
gg3 = plotLearnerPrediction("classif.naiveBayes", task = iris.task, features = c("Sepal.Length", "Sepal.Width"))

grid.arrange(gg1, gg2, gg3, ncol = 1)
```
