## Data and Tasks

### Study Goals

*Theoretical (T)*

- Be able to differentiate between regression and classification tasks
- Understand what the data generating process is.

*Practical (P)*

- Get to know how to create tasks in `mlr3`

### Preparation

1.  *(T)* Watch the following videos:
    <center>
    ![Tasks and Data](https://youtu.be/d9YbO6P4AdU){width="75%"}
    </center>


<!-- TODO: adapt to mlr3 -->
1.  *(P)* Read the [`mlr3` tutorial about tasks](https://mlr3book.mlr-org.com/tasks.html). 


### Exercises

#### *(T)* Quiz

```{r tasks-quiz, echo=FALSE}
question("Which statements are true?",
    answer("Classification is a supervised learning task.",
        correct = TRUE),
    answer("Regression is a supervised learning task.",
        correct = TRUE),
    answer("Clustering is a supervised learning task."),
    answer("We assume that a probability distribution characterizes the process that generates the observed data. This is called the *data generating process*.", correct = TRUE),
    answer("The data generating process is the true underlying phenomenon
  creating the data.", correct = TRUE)
)
```

#### *(P)* Creating tasks with `mlr3`

For this exercise, take a look at the [`iris` dataset](https://en.wikipedia.org/wiki/Iris_flower_data_set):

```{r}
summary(iris)
```

1.  Define a new classification task called `iris_task` based on the `iris` dataset with target variable `Species` and just two features `Sepal.Width` and `Petal.Width`:


```{r iristask, exercise=TRUE, exercise.lines=5, exercise.checker=taskChecker("iris_task")}
iris_task <- 
```

```{r iristask-hint-1}
# Use the 'TaskClassif$new()' function of mlr3.
# The identifier for the task 'id' can be arbitrarily chosen, but must be set.

iris_task <- TaskClassif$new(id = ..., 
                backend = ..., 
                target = ...)
```

```{r iristask-hint-2}
# Select just the three columns Sepal.Width, Petal.Width, and Species
iris[, c("Species", "Sepal.Width", "Petal.Width")]
```

```{r iristask-hint-3}
# The target variable has to be specified as character of length 1
"Species"
```

```{r iristask-solution}
iris_task <- TaskClassif$new(id = "iris_task", 
                backend = iris[, c("Species", "Sepal.Width", "Petal.Width")], 
                target = "Species")
```

```{r iristask-check}
iris_task <- TaskClassif$new(id = "iris_task", 
                backend = iris[, c("Species", "Sepal.Width", "Petal.Width")], 
                target = "Species")
```

