## Data and Tasks

### Study Goals

*Theoretical (T)*

- Be able do differentiate between regression and classification tasks

*Practical (P)*

- Get to know how to create tasks in `mlr`

### Preparation

1.  *(T)* Watch the following videos:
    <center>
    ![Tasks and Data](https://youtu.be/d9YbO6P4AdU){width="75%"}
    </center>


1.  *(P)* Read the [`mlr` tutorial about tasks](https://mlr.mlr-org.com/articles/tutorial/task.html).

### Exercises

#### *(T)* Quiz

```{r tasks-quiz, echo=FALSE}
# quiz(caption = "Tasks",
  question("Which statements are true?",
    answer("Supervised learning tasks are prediction problems.",
           correct = TRUE),
    answer("Unsupervised learning tries to discover structure and patterns in
           the training data.",
           correct = TRUE),
    answer("Every unsupervised task is internally treated as supervised task."),
    answer("Unsupervised learning is learning without a target variable.",
           correct = TRUE),
    answer("Classification is a supervised learning task.",
           correct = TRUE),
    answer("Regression is a supervised learning task.",
           correct = TRUE),
    answer("Clustering is a supervised learning task.")
  )
```

#### *(P)* Creating tasks with `mlr`

For this exercise, take a look at the [`iris` dataset](https://en.wikipedia.org/wiki/Iris_flower_data_set):

```{r}
summary(iris)
```

1.  Define a new classification task called `iris_task` based on the `iris` dataset with target variable `Species` and just two features `Sepal.Width` and `Petal.Width`:

```{r iristask, exercise=TRUE, exercise.lines=5, exercise.checker=taskChecker("iris_task")}
iris_task =
```

```{r iristask-hint-1}
# Use the 'makeClassifTask' function of mlr
makeClassifTask(...)
```

```{r iristask-hint-2}
# Select just the three columns Sepal.Width, Petal.Width, and Species
iris[, c("Species", "Sepal.Width", "Petal.Width")]
```

```{r iristask-hint-3}
# The target variable has to be specified as character of length 1
"Species"
```

```{r iristask-solution}
iris_task = makeClassifTask(data = iris[, c("Species", "Sepal.Width", "Petal.Width")], target = "Species")
```

```{r iristask-check}
iris_task = makeClassifTask(data = iris[, c("Species", "Sepal.Width", "Petal.Width")], target = "Species")
```

