## Tasks

```{r, include=FALSE}
taskChecker = function (label, user_code, check_code, envir_result, evaluate_result, ...)
{
  print(user_code)
  setup_state(sol_code = check_code, stu_code = user_code)

  msg = errorToMessage(expr = {
    # checkRhs(user_code)
    ex() %>% check_object("iris_task") %>% check_equal()
  })
  if (! is.null(msg))
    return(msg)

  return(list(message = "Great job! :)", correct = TRUE, location = "append"))
}
```

**Study Goals**

- Be able do differentiate between regression and classification tasks
- Get to know how to create tasks in `mlr`

**Preparation**

1.  Watch the following video  (sorry, rather low volume...):
    <center>
    ![](https://www.youtube.com/watch?v=pFWVIGPpupk&t=0s&index=3&list=PLMyWaJl2LoXyhFvMjtbBGs0Pi8khHbKm3){width="75%"}
    </center>

1.  Read the [`mlr` tutorial about tasks](https://mlr.mlr-org.com/articles/tutorial/task.html).

**Exercises**

```{r tasks-quiz, echo=FALSE}
# quiz(caption = "Tasks",
  question("Which statements are true?",
    answer("Supervised learning tasks are prediction problems.",
           correct = TRUE),
    answer("Unsupervised learning tries to discover structure and patterns in 
           the training data.",
           correct = TRUE),
    answer("Every unsupervised task is internally treated as supervised task."),
    answer("Unsupervised learning is learning without a target variable.", 
           correct = TRUE),
    answer("Classification is a supervised leaning task.", 
           correct = TRUE),
    answer("Regression is a supervised learning task.", 
           correct = TRUE),
    answer("Clustering is a supervised learning task.")
  )
```

### Creating tasks with `mlr`

For this exercise, take a look at the [`iris` dataset](https://en.wikipedia.org/wiki/Iris_flower_data_set):

```{r}
summary(iris)
```

1.  Define a new classification task based on the iris dataset with target variable `Species` and just two features `Sepal.Width` and `Petal.Width`:

```{r iristask, exercise=TRUE, exercise.lines=5, exercise.checker=taskChecker}
iris_task =
```

```{r iristask-hint-1}
# Use the 'makeClassifTask' function of mlr
makeClassifTask(...)
```

```{r iristask-hint-2}
# Select just the three columns Sepal.Width, Petal.Width, and Species
iris[, c("Species", "Sepal.Width", "Petal.Width")]
```

```{r iristask-hint-3}
# The target variable has to be specified as character of length 1
"Species"
```

```{r iristask-solution}
makeClassifTask(data = iris[, c("Species", "Sepal.Width", "Petal.Width")], target = "Species")
```

```{r iristask-check}
iris_task = makeClassifTask(data = iris[, c("Species", "Sepal.Width", "Petal.Width")], target = "Species")
```

