## Linear Regression

```{r, include=FALSE}
data(BostonHousing, package = "mlbench")
regr.task = makeRegrTask(id = "bh", data = BostonHousing, target = "medv")
```

**Study Goals**

- Learn how linear regression model works
- Get to know how to fit a linear regression model in `mlr`

**Preparation**

1. Watch the following video  (sorry, rather low volume...):
<center>
![](https://www.youtube.com/watch?v=v9eocCAfJ_w&list=PLMyWaJl2LoXyhFvMjtbBGs0Pi8khHbKm3&index=6&t=0s){width="75%"}
</center>

2. Read the [`mlr` tutorial about learners](https://mlr.mlr-org.com/articles/tutorial/learner.html) and how to [train](https://mlr.mlr-org.com/articles/tutorial/train.html) them.

### Create the `mlr` learner

Create a linear regression learner with `predict.type = 'se'` :
  
```{r lm-definition, exercise=TRUE, exercise.lines=5, exercise.checker=checker}
lm_learner =
```

```{r lm-definition-check}
lm_learner = makeLearner("regr.lm", predict.type = 'se')
```

### Train the `mlr` learner

Now train the learner on the task defined in session 2:
  
```{r lm-training, exercise=TRUE, exercise.lines=5, exercise.checker=checker}
mod.lm =
```

```{r, include=FALSE}
lm_learner = makeLearner("regr.lm", predict.type = 'se')
```

```{r lm-training-check}
n = getTaskSize(regr.task)
# Use 1/4 of the observations for training
train.set = sample(n, size = n/4)
mod.lm = train(lm_learner, regr.task, subset = train.set)

```
