## Tasks

**Study Goals**

- Be able do differentiate between regression and classification tasks
- Get to know how to create tasks in `mlr`

**Preparation**

1.  Watch the following video:
    <center>
    ![](https://www.youtube.com/watch?v=pFWVIGPpupk){width="50%"}
    </center>

1.  Read the [`mlr` tutorial about tasks](https://mlr.mlr-org.com/articles/tutorial/task.html)

### Creating tasks with `mlr`

For this exercise, take a look at the [`iris` dataset](https://en.wikipedia.org/wiki/Iris_flower_data_set):

```{r}
summary(iris)
```

1.  Define a new classification task based on the iris dataset:

```{r iristask, exercise=TRUE, exercise.lines=5, exercise.checker=checker}
iris_task =
```

```{r iristask-check}
iris_task = makeClassifTask(data = iris, target = "Species")
```

2. Now, try to define a regression task on iris

```{r falseiristask, exercise=TRUE, exercise.lines=5, exercise.checker=checker}
iris_task =
```

```{r falseiristask-check}
iris_task = makeRegrTask(data = iris, target = "Species")
```

