## Discussion

### Study Goals

*Theoretical (T)*

- Understand the pros and cons of trees


*Practical (P)*

- Train a model for classification or regression with CART using `mlr3`
- Visualize the tree structure and decision boundaries for different data situations


### Preparation

1.  *(T)* Watch the following video:
    <center>
    ![Discussion](https://youtu.be/nKULLVAUk74){width="75%"}
    </center>

1.  *(P)* Be sure to know how to define tasks and learners in `mlr3`. If not, take a look at the tutorials about these topics.

### Exercises

#### *(T)* Quiz

```{r tree-quiz-discussion, echo=FALSE}
question("Which statements about CART are true?",
  answer("CART is a stable algorithm, if the data changes slightly the tree structure remains quite similar."),
  answer("CART is robust to outliers.", correct = TRUE),
  answer("With trees it is easy to handle missing values.", correct = TRUE),
  answer("A simple tree is everything you need to fit a linear function $y \\approx a + b*x$ with numeric target.")
)
```

#### *(P)* Create the spiral task

For this exercise we are, again, using simulated data. Here we use the package `mlbench` to simulate spiral data. Therefore, use the function `mlbench.spirals()` to generate 500 data points with `sd = 0.1` and store it into a object called `spirals`. Transform this object to a `data.frame` and define the task with `target = "classes"`:

```{r spiral-task, exercise=TRUE, exercise.checker=taskChecker("spirals_task")}
library(mlbench)
library(ggplot2)

set.seed(314)
spirals <-
spirals <-
spirals_task <-

# Visualization of the data
ggplot(data = spirals, aes(x.1, x.2, color = classes)) + geom_point()
```

```{r spiral-task-hint-1}
# Generate data points and transfrom them to a data.frame
set.seed(314)
spirals <- mlbench.spirals(500, sd = 0.1)
spirals <- as.data.frame(spirals)
```

```{r spiral-task-hint-2}
# Define the task
spirals_task <- TaskClassif$new(id = "spirals_task", backend = spirals, target = "classes")
```

```{r spiral-task-solution}
library(mlbench)
library(ggplot2)

set.seed(314)
spirals <- mlbench.spirals(500, sd = 0.1)
spirals <- as.data.frame(spirals)
spirals_task <- TaskClassif$new(id = "spirals_task", backend = spirals, target = "classes")

# Visualization of the data
ggplot(data = spirals, aes(x.1, x.2, color = classes)) + geom_point()
```

```{r spiral-task-check}
library(mlbench)
library(ggplot2)

set.seed(314)
spirals <- mlbench.spirals(500, sd = 0.1)
spirals <- as.data.frame(spirals)
spirals_task <- TaskClassif$new(id = "spirals_task", backend = spirals, target = "classes")

# Visualization of the data
ggplot(data = spirals, aes(x.1, x.2, color = classes)) + geom_point()
```


#### *(P)* Create and train the `mlr3` CART learner

Use the `classif.rpart` learner to train the tree. Use `minsplit = 20`  and `cp = 0.01` as hyperparameters:

```{r cart-model, exercise=TRUE, exercise.checker=modelChecker("learner_cart")}
library(mlbench)

set.seed(314)
spirals <-
spirals <-
spirals_task <-

learner_cart <-
learner_cart$
```

```{r cart-model-hint-1}
# Use the objects defined previously
set.seed(314)
spirals <- mlbench.spirals(500, sd = 0.1)
spirals <- as.data.frame(spirals)
spirals_task <- TaskClassif$new(id = "spirals_task", backend = spirals, target = "classes")
```

```{r cart-model-hint-2}
# To define the learner use 'lrn()'. Hyperparameter are passed as comma separated argument

learner_cart <- lrn("classif.rpart", minsplit = 20, cp = 0.01)
```

```{r cart-model-hint-3}
# Finally, use 'train()' to calculate the model
learner_cart$train(learner_cart, spirals_task)
```

```{r cart-model-solution}
library(mlbench)

set.seed(314)
spirals <- mlbench.spirals(500, sd = 0.1)
spirals <- as.data.frame(spirals)
spirals_task <- TaskClassif$new(id = "spirals_task", backend = spirals, target = "classes")

learner_cart <- lrn("classif.rpart", minsplit = 20, cp = 0.01)
learner_cart$train(spirals_task)
```

```{r, cart-model-check}
library(mlbench)

set.seed(314)
spirals <- mlbench.spirals(500, sd = 0.1)
spirals <- as.data.frame(spirals)
spirals_task <- TaskClassif$new(id = "spirals_task", backend = spirals, target = "classes")

learner_cart <- lrn("classif.rpart", minsplit = 20, cp = 0.01)
learner_cart$train(spirals_task)
```

#### *(P)* Visualize the fitted tree

You can use the function `rpart.plot()` from the package of the same name `rpart.plot` to visualize the tree structure. For this, you have to access the internal fitted model which is stored as `state$model` element of the fitted model. Additionally, call `plot_learner_prediction()` to visualize the decision boundaries:

```{r cart-viz, exercise=TRUE}
library(mlbench)
library(rpart.plot)

set.seed(314)
spirals <-
spirals <-
spirals_task <-

learner_cart <-
learner_cart$

rpart.plot(..., roundint = FALSE)
plot_learner_prediction(learner = ..., task = ...)
```

```{r cart-viz-hint-1}
# Use the objects defined previously and library 'rpart.plot'
library(mlbench)
library(rpart.plot)

set.seed(314)
spirals <- mlbench.spirals(500, sd = 0.1)
spirals <- as.data.frame(spirals)
spirals_task <- TaskClassif$new(id = "spirals_task", backend = spirals, target = "classes")

learner_cart <- lrn("classif.rpart", minsplit = 20, cp = 0.01)
learner_cart$train(spirals_task)
```

```{r cart-viz-hint-2}
# Call 'rpart.plot' of the fitted rpart model. You can access the model via the 'learner.model' element of the model
rpart.plot(learner_cart$state$model, roundint = FALSE)
```

```{r cart-viz-hint-3}
# Call 'plotLearnerPrediction()' on the learner and task to get decision boundaries
plot_learner_prediction(learner = learner_cart, task = spirals_task)
```

```{r cart-viz-solution}
library(mlbench)
library(rpart.plot)

set.seed(314)
spirals <- mlbench.spirals(500, sd = 0.1)
spirals <- as.data.frame(spirals)
spirals_task <- TaskClassif$new(id = "spirals_task", backend = spirals, target = "classes")

learner_cart <- lrn("classif.rpart", minsplit = 20, cp = 0.01)
learner_cart$train(spirals_task)

rpart.plot(learner_cart$state$model, roundint = FALSE)
plot_learner_prediction(learner = learner_cart, task = spirals_task)
```

#### *(P)* Visualize the tree for different data constellations

To see how trees behave on the same data situation but different data constellation draw 500 rows randomly (using the `sample()` function) with replacement and recreate the figures:

```{r cart-viz-data-setup}
library(mlbench)
library(rpart.plot)

set.seed(314)
spirals = mlbench.spirals(500, sd = 0.1)
spirals = as.data.frame(spirals)

set.seed(Sys.time())
```

```{r cart-viz-data, exercise=TRUE}
# Choose different observations for training the model
train_idx <- sample(...)

spirals_task <- TaskClassif$new(id = "spirals_task", backend = spirals, target = "classes")

learner_cart <- lrn("classif.rpart", minsplit = 20, cp = 0.01)
learner_cart$train(spirals_task)

rpart.plot(learner_cart$state$model, roundint = FALSE)
plot_learner_prediction(learner = learner_cart, task = spirals_task)
```

```{r cart-viz-data-hint-1}
# Use different data situation, e.g. via sampling:
train_idx <- sample(1:500, size = 500, replace = TRUE)
```

```{r cart-viz-data-solution}
# Choose different observations for training the model
train_idx <- sample(1:500, size = 500, replace = TRUE)

spirals_task <- TaskClassif$new(id = "spirals_task", backend = spirals, target = "classes")

learner_cart <- lrn("classif.rpart", minsplit = 20, cp = 0.01)
learner_cart$train(spirals_task)

rpart.plot(learner_cart$state$model, roundint = FALSE)
plot_learner_prediction(learner = learner_cart, task = spirals_task)
```

```{r cart-viz-data-quiz, echo=FALSE}
question("Which statements are true?",
  answer("The data situation is still the same, we are just using less unique data points.", correct = TRUE),
  answer("The tree structure for different data situations remains the same."),
  answer("The decision boundaries for different data situations remain the same.")
)
```


#### *(P)* Visualize the tree for different hyperparameters

Finally, visualize the tree structure and decision boundaries for varying hyperparameters. Also try to understand how changes in the parameters affect the figures.

```{r cart-viz-hyperpars-setup}
library(mlbench)
library(rpart.plot)

set.seed(314)
spirals <- mlbench.spirals(500, sd = 0.1)
spirals <- as.data.frame(spirals)
spirals_task <- TaskClassif$new(id = "spirals_task", backend = spirals, target = "classes")
```

```{r cart-viz-hyperpars, exercise=TRUE}
# Choose different hyperparameter configurations
minsplit <- 20
cp <- 0.01

learner_cart <- lrn("classif.rpart", minsplit = minsplit, cp = cp)
learner_cart$train(spirals_task)

rpart.plot(learner_cart$state$model, roundint = FALSE)
plot_learner_prediction(learner = learner_cart, task = spirals_task)
```

```{r cart-viz-hyperpars-quiz, echo=FALSE}
# mlr3 does not support minbucket as a parameter at the moment

question("Which statements are true?",
  answer("Reducing just the complexity `cp` leads to bigger trees."),
  answer("Reducing complexity `cp`, minsplit at the same time leads to bigger trees.", correct = TRUE),
  #answer("Reducing complexity `cp`, minsplit, and minbucket at the same time leads to bigger trees.", correct = TRUE),
  #answer("It doesn't make sense to choose minsplit bigger than minbucket.", correct = TRUE),
  answer("A very small minsplit leads to smaller trees."),
  #answer("A very small minbucket leads to smaller trees."),
  answer("The bigger the tree the more unstable its structure becomes.", correct = TRUE)
)
```